---
title: "Peer assignment 1 of reproducible Research"
author: "Jie Yong"
date: "Tuesday, December 09, 2014"
output: html_document
---

###1. Loading and preprocessing the data:  
read the csv file and change the type of the date column

```{r loading}
data <-read.csv("C:/R/activity.csv")
data$date <-as.Date(data$date)
```

###2. What is mean total number of steps taken per day?
First calculate the sum of the steps for each day then make a histogram. Next I calculate the mean and median of the steps for each day.  
```{r mean and median}
sum <-tapply(data$steps, data$date, sum)
hist(sum,breaks=20)
dev.copy(png,file="instructions_fig/hist1.png")
dev.off()
mean0 <- as.character(round(mean(sum,na.rm=TRUE),2))
median0 <-as.character(round(median(sum, na.rm=TRUE),2))

```
total number of steps taken per day: the mean is `r mean0`. The median is `r median0`

###3. What is the average daily activity pattern?
```{r daily}
library(lattice)
data$daily <-tapply(data$steps, data$interval,mean,na.rm=TRUE)
xyplot(daily ~ interval, data=data, type="l")
dev.copy(png,file="instructions_fig/daily_average.png")
dev.off()
inx <- which.max(data$daily)
inx
```
The interval `r inx` has the maximum of average daily activity


###4. Imputing missing values  
the missing data is replaced by the daily average of that interval.
```{r}
row.has.na <- apply(data, 1, function(x){any(is.na(x))})
#total number of rows that have NAs are:
sum(row.has.na)

dat <- data
daily <-tapply(data$steps, data$interval,mean,na.rm=TRUE)

for (i in which(is.na(data$steps))) {
  index <-data[i,3]/5+1
  dat[i,1] <- round(daily[index],2)
}

sum2 <-tapply(dat$steps, dat$date, sum)
hist(sum2,breaks=20)
dev.copy(png,file="instructions_fig/hist2.png")
dev.off()
mean2 <- as.character(round(mean(sum2,na.rm=TRUE),2))
median2 <-as.character(round(median(sum2, na.rm=TRUE),2))
```
After imputing the missing values, the new mean is `r mean2`. The new median is `r median2`. Their values do not change.


###5. Are there differences in activity patterns between weekdays and weekends?
```{r}
dat$date <-as.factor(ifelse(weekdays(dat$date) %in% c("Saturday","Sunday"), "Weekend", "Weekday"))
result <- aggregate(steps ~ date + interval, data = dat, FUN= "mean" )
xyplot(steps~interval|date,data=result,type="l",layout=c(1,2))
dev.copy(png,file="instructions_fig/difference.png")
dev.off()
```


As you can see, there is a significant difference between weekday and weekend activity patterns.

The end.
